Thermodynamic response functions in a cell fluid model

### 2024-08-14 main_morse_response_functions.tex
add formulas for density and average number of particles
made refinement to eos via dens and temp

	The function $E_\mu(M, T)$ from the equation \eqref{eq:eosMT} is provided by
	\begin{align}
		& E_\mu (M, T) = - \frac{\ln (2\pi \beta W(0))}{2 N_v}  +  g_0 - \frac{\beta W(0)}{2} \!\! \left( \! \frac{\tilde\mu}{W(0)} \! \right)^{\! 2} \!\!\! - \frac{g_3}{g_4} {M} \! - \frac{g_3^2}{2 g_4^2}  \tilde D(0) - \frac{1}{24} \frac{g_3^4}{g_4^3}. 
	\end{align}
	Here the quantity $N_v$ defines the number of cubic cells in volume $V$ for the initial model.
	In the thermodynamic limit, $N_v \to \infty$.

@@

	The density of the many-particle system is the ratio of the average number of particles $\left\langle N \right\rangle$  to the total volume of the system $V$.
	In the framework of the grand canonical ensemble the average number of particles $\left\langle N \right\rangle$ is a partial derivative of natural logarithm of the grand partition function $\Xi$ with respect to the product $\beta \mu$. Using the general relation for the equation of state $PV\beta = \ln \Xi $ the density reads
	\begin{equation}
	\eta = \frac{\langle N\rangle}{v N_v} = \frac{1}{v} \frac{\partial}{\partial\beta\mu} Pv\beta = \frac{\bar n}{v},
	\end{equation} 
	where	
	\begin{equation}\label{eq:density}
	\bar n = n_g - \bar{M} + \frac{ \bar \rho_0}{\beta W(0)}.
	\end{equation}
	The quantity $n_g$ in the equation \eqref{eq:density} is the critical density
	\begin{equation}\label{eq:crit_dens}
	n_g = g_1 - \frac{g_3}{g_4}\left(  g_2 - \frac{1}{2} \frac{g_3^2}{g_4}\right) - \frac{1}{6} \frac{g_3^3}{g_4^2}.
	\end{equation}
	Combination of the equations \eqref{eq:density} and \eqref{eq:ro_M} reveals the explicit expression for the effective chemical potential $M$ as a function of the density and the temperature
	\begin{equation}\label{eq:M_nT}
		\bar M = \frac{\rho_{n}}{\beta W(0)} - (\bar n - n_g),
	\end{equation}

	********************************************************

	[THE FOLLOWING NEEDS REFINEMENT, OKSANA WILL DO IT SOON]

	where
	\begin{equation} \label{eq:ro_nT}
		 \rho_{n} = - 2 \left(\frac{-2\left(  g_2 - \frac{1}{2} \frac{g_3^2}{g_4}\right)}{g_4}\right)^{\frac{1}{2}} \cos \left( \frac{\alpha_n}{3} + \frac{\pi}{3} \right), \qquad 	\alpha_n = \arccos \left( \frac{3}{2} \left( - \frac{g_4}{2\left(  g_2 - \frac{1}{2} \frac{g_3^2}{g_4}\right)^3}\right)^{\frac{1}{2}} (n_g - \bar n)\right). 
	\end{equation}

	At $T>T_c$ the equation of state of a cell fluid model in terms of density has the following form
	\begin{equation}\label{eq:eosNT}
	Pv\beta = E_\mu^{(2)}(n,T) + \bar M \rho_{n} + \frac{\tilde D(0)}{2} \rho_{n}^2 - \frac{a_4}{24} \rho_{n}^4,
	\end{equation}
	where
	\begin{equation}\label{3d30}
	\bar M = \frac{\rho_{n}}{\beta W(0)} - (\bar n - n_g)
	\end{equation}
	is a function of density and temperature. Recall the formula \eqref{2d22} to find
	\begin{equation}\label{3d31}
	\frac{\tilde\mu}{W(0)} = \bar M - n_g + n_c \tilde g_2 \gamma_\tau,
	\end{equation}
	where $n_g$ is defined in \eqref{3d4}. Therefore,
	\begin{align}\label{3d32}
		E_\eta (\bar n,T) & = - \frac{\ln (2\pi \beta W(0))}{2 N_v} + g_0 - \frac{1}{2\tilde g_2 \gamma_\tau} \left( \frac{\tilde\mu}{W(0)}\right)  + n_c \bar M- \nonumber \\
		& - \frac{1}{2} n_c^2 \tilde D(0) - \frac{1}{24} \frac{g_3^4}{g_4^3}.
	\end{align}

	********************************************************

>>>>

	The function $E_\mu(M, T)$ from the equation \eqref{eq:eosMT} is provided by
	\begin{align}\label{eq:E_mu}
		& E_\mu (M, T) = - \frac{\ln (2\pi \beta W(0))}{2 N_v}  +  g_0 - \frac{\beta W(0)}{2} \!\! \left( \! \frac{\tilde\mu}{W(0)} \! \right)^{\! 2} \!\!\! - \frac{g_3}{g_4} {M} \! - \frac{g_3^2}{2 g_4^2}  \tilde D(0) - \frac{1}{24} \frac{g_3^4}{g_4^3}. 
	\end{align}
	Here the quantity $N_v$ defines the number of cubic cells in volume $V$ for the initial model.
	In the thermodynamic limit, $N_v \to \infty$.

@@

	The density $\eta$ of the many-particle system is the ratio of the average number of particles $\left\langle N \right\rangle$  to the total volume of the system $V$.
	In the framework of the grand canonical ensemble the average number of particles $\left\langle N \right\rangle$ is a partial derivative of natural logarithm of the grand partition function $\Xi$ with respect to the product $\beta \mu$. Using the general relation for the equation of state $PV\beta = \ln \Xi $ the density reads
	\begin{equation}
	\eta = \frac{\langle N\rangle}{v N_v} = \frac{1}{v} \frac{\partial}{\partial\beta\mu} Pv\beta = \frac{\bar n}{v},
	\end{equation} 
	where	
	\begin{equation}\label{eq:density}
	\bar n = n_g - \bar{M} + \frac{ \bar \rho_0}{\beta W(0)}.
	\end{equation}
	The quantity $n_g$ in the equation \eqref{eq:density} is the critical density
	\begin{equation}\label{eq:crit_dens}
	n_g = g_1 - \frac{g_3}{g_4}\left(  g_2 - \frac{1}{2} \frac{g_3^2}{g_4}\right) - \frac{1}{6} \frac{g_3^3}{g_4^2}.
	\end{equation}
	Combination of the equations \eqref{eq:density} and \eqref{eq:ro_M} reveals the explicit expression for the effective chemical potential $M$ as a function of the density and the temperature
	\begin{equation}\label{eq:M_nT}
		\bar M = \frac{\rho_{n}}{\beta W(0)} - (\bar n - n_g),
	\end{equation}
	where
	\begin{align} \label{eq:ro_nT}
		 & \rho_{n} = - 2 \left(\frac{g_3^2 - 2g_2 g_4}{g_4^2} \right)^{\frac{1}{2}} \cos \left( \frac{\alpha_n}{3} + \frac{\pi}{3} \right), \\
		 & \alpha_n = \arccos \left( \frac{3}{2} \left( - \frac{g_4^4}{\left( 2 g_2 g_4 - g_3^2\right)^3}\right)^{\frac{1}{2}} (n_g - \bar n)\right). \nonumber
	\end{align}

	At $T>T_c$ the equation of state of a cell fluid model in terms of the density and the temperature has the following form
	\begin{equation}\label{eq:eosNT}
	Pv\beta = E_\eta (\bar n,T) + \bar M \rho_{n} + \frac{\tilde D(0)}{2} \rho_{n}^2 - \frac{a_4}{24} \rho_{n}^4,
	\end{equation}
 The function $E_\nu (\bar n, T)$ in \eqref{eq:ro_nT} is the function $E_\mu (M,T)$ \eqref{eq:E_mu} rewritten in terms of density and temperature taking into account the expression \eqref{eq:M_nT}.
